---
title: "MATH 216 Homework 1"
author: "Alexander Pastora"
output:
  html_document:
    collapsed: no
    smooth_scroll: no
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
# For read_csv() command, which is an improved version of base R's read.csv()
library(readr) 

# Load data sets. Note this assumes this .Rmd files is in the same directory as
# the .csv files.
flights <- read_csv("data/flights.csv") %>% 
  mutate(date=as.Date(date))
weather <- read_csv("data/weather.csv") %>% 
  mutate(date=as.Date(date))
planes <- read_csv("data/planes.csv")
airports <- read_csv("data/airports.csv")
states <- read_csv("data/states.csv")
```





## Admistrative:

Please indicate

* Who you collaborated with:
* Roughly how much time you spent on this HW so far:
* The URL of the RPubs published URL [here](https://www.nhl.com/).
* What gave you the most trouble:
* Any comments you have:





## Question 1:

Plot a "time series" of the proportion of flights that were delayed by > 30 minutes on each day.  i.e.
 
* the x-axis should be some notion of time
* the y-axis should be the proportion.

Using this plot, indicate describe the
[seasonality](https://en.wikipedia.org/wiki/Seasonality) of when delays over 30
minutes tend to occur.

```{r, echo=FALSE, message=FALSE, fig.width=12, fig.height=6}
#X-axis should be either date, or hour. Must calculate delays greater than 30 min.

flight_delays <- flights %>% filter(!is.na(dep_delay)) %>% select(dep,dep_delay)

# Calculate proportion
big_delays <- flight_delays %>% group_by(dep) %>% tally(dep_delay > 30)
total_delays <- flight_delays %>% group_by(dep) %>% tally()
delays_by_hour <- inner_join(big_delays,total_delays, by="dep")
delays_by_hour <-rename(delays_by_hour, big_delays=n.x)
delays_by_hour <-rename(delays_by_hour, total_flight=n.y)
delays_by_hour <- mutate(delays_by_hour, prop = big_delays/total_flight)

#Graph everything
ggplot(data=delays_by_hour, aes(x=dep, y=prop)) +geom_bar(stat="identity") + geom_smooth(se=FALSE)
```

As time of departure increases, the proportion of flights with major delays (greater than 30 minutes) increases as well. For the purposes of this data, a "day" starts at around 6AM, and ends at about 2AM.



## Question 2:

Some people prefer flying on older planes.  Even though they aren't as nice,
they tend to have more room.  Which airlines should these people favor?

```{r, echo=FALSE, fig.width=12, fig.height=6}

```





## Question 3:

* What states did Southwest Airlines' **flight paths** tend to fly to?
* What states did Southwest Airlines' **flights** tend to fly to?

For example, Southwest Airlines Flight 60 to Dallas consists of a single flight
path, but since it flew 299 times in 2013, it would be counted as 299 flights.

```{r, echo=FALSE, message=FALSE, fig.width=12, fig.height=6}

```





## Question 4:

I want to know proportionately what regions (NE, south, west, midwest) each 
carrier flies to/from Houston in the month of July.  Consider the `month()`
function from the `lubridate` package.

```{r, echo=FALSE, message=FALSE, fig.width=12, fig.height=6}

```
