---
title: "MATH 216 Homework 1"
author: "Alexander Pastora"
output:
  html_document:
    collapsed: no
    smooth_scroll: no
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
# For read_csv() command, which is an improved version of base R's read.csv()
library(readr) 
library(lubridate)

# Load data sets. Note this assumes this .Rmd files is in the same directory as
# the .csv files.
flights <- read_csv("data/flights.csv") %>% 
  mutate(date=as.Date(date))
weather <- read_csv("data/weather.csv") %>% 
  mutate(date=as.Date(date))
planes <- read_csv("data/planes.csv")
airports <- read_csv("data/airports.csv")
states <- read_csv("data/states.csv")
```





## Admistrative:

Please indicate

* Who you collaborated with: Nobody
* Roughly how much time you spent on this HW so far: 5 hours
* The URL of the RPubs published URL [here](http://rpubs.com/pinkfan9999/Pre-HW1).
* What gave you the most trouble: Trying to get rid of the duplicate planes in question number 2, and trying to isolate the July flights in question four.
* Any comments you have:





## Question 1:

Plot a "time series" of the proportion of flights that were delayed by > 30 minutes on each day.  i.e.
 
* the x-axis should be some notion of time
* the y-axis should be the proportion.

Using this plot, indicate describe the
[seasonality](https://en.wikipedia.org/wiki/Seasonality) of when delays over 30
minutes tend to occur.

```{r, echo=FALSE, message=FALSE, fig.width=12, fig.height=6}
#X-axis should be either date, or hour. Must calculate delays greater than 30 min.

flight_delays <- flights %>% filter(!is.na(dep_delay)) %>% select(date,dep_delay)

# Calculate proportion
big_delays <- flight_delays %>% group_by(date) %>% tally(dep_delay > 30)
total_delays <- flight_delays %>% group_by(date) %>% tally()
delays_by_date <- inner_join(big_delays,total_delays, by="date") %>% 
  rename(big_delays=n.x) %>% 
  rename(total_flight=n.y) %>% 
  mutate(prop = big_delays/total_flight)

#Graph everything
ggplot(data=delays_by_date, aes(x=date, y=prop)) +geom_bar(stat="identity") + geom_smooth(se=FALSE) + xlab("Date") + ylab("Proportion of Flights Delayed > 30 minutes")
```

The change in proportion of flights delayed more than 30 minutes seems to be fairly cyclical. From January 2011 to July 2011, the proportion increases. After July 2011, the proportion decreases until October 2011. The proportion then begins to increase onwards. 


## Question 2:

Some people prefer flying on older planes.  Even though they aren't as nice,
they tend to have more room.  Which airlines should these people favor?

```{r, echo=FALSE, fig.width=12, fig.height=6}
# Get rid of missing years
planes <- planes %>% filter(!is.na(year))

# Adjust flight data in a new data table
flight_data <- flights %>% group_by(plane) %>% 
  select(plane,carrier) %>%
  mutate(duplicate = duplicated(plane)) %>% 
# Get rid of Duplicate planes
  filter(duplicate =="FALSE")

old_planes <- left_join(planes, flight_data, by= "plane")%>% 
  group_by(plane) %>% 
  select(plane,carrier,year) %>% 
  mutate(plane_age = 2016-year)

ggplot(data=old_planes, aes(x=carrier, y=plane_age)) + geom_boxplot()+ xlab("Airline Carrier") +ylab("Plane Age (Years)") 
```

People wanting to fly on old planes should fly with MQ (Envoy Air) or AA (American Airlines), because these airlines tend to have a good distribution of older planes. On a side note, Envoy Air is one of American Airline's regional airlines operating under American Eagle.



## Question 3:

* What states did Southwest Airlines' **flight paths** tend to fly to?
* What states did Southwest Airlines' **flights** tend to fly to?

For example, Southwest Airlines Flight 60 to Dallas consists of a single flight
path, but since it flew 299 times in 2013, it would be counted as 299 flights.

```{r, echo=FALSE, message=FALSE, fig.width=12, fig.height=6}
southwest_flights <- flights %>% filter(carrier=="WN") %>% select(carrier,flight, dest, time)

southwest_flights <- left_join(southwest_flights,airports, by=c("dest"="iata"))%>% 
  filter(!is.na(state))

southwest_flightpath <- southwest_flights %>% 
  group_by(flight) %>% 
  mutate(duplicate_flight = duplicated(dest)) %>% 
  filter(duplicate_flight==FALSE) %>% 
  group_by(state) %>% 
  tally() %>% 
  rename(flight_paths = n)
ggplot(southwest_flightpath, aes(x=state, y=flight_paths)) +geom_bar(stat="identity") + ggtitle("Southwest Flight Paths") +xlab("State") + ylab("Number of Flight Paths")

southwest_flight_total <- southwest_flights %>% group_by(state) %>% 
  tally(time)%>% 
  rename(flights=n)
ggplot(southwest_flight_total, aes(x=state, y=flights)) +geom_bar(stat="identity") +ggtitle("Southwest Flights") +xlab("State") + ylab("Number of Flights")

```

According to the data, the most flight paths, for Southwest Airlines, were to Texas, Florida, Louisiana, California, and Okhlahoma. 

However, the most flights were to Texas, California, Florida, Nevada, and Illinois.


## Question 4:

I want to know proportionately what regions (NE, south, west, midwest) each 
carrier flies to/from Houston in the month of July.  Consider the `month()`
function from the `lubridate` package.

```{r, echo=FALSE, message=FALSE, fig.width=12, fig.height=6}

July_flights <- flights %>% filter(date >= ymd(20110701) & date <= ymd(20110731))
July_flights <- left_join(July_flights,airports, by=c("dest"="iata"))
July_flights <- left_join(July_flights, states, by="state")

carrier_flights <-July_flights %>% filter(!is.na(region)) %>%  group_by(carrier, region)  %>% tally()

total <- July_flights %>% filter(!is.na(region)) %>% group_by(carrier) %>% tally()

carrier_flights <- inner_join(carrier_flights, total, by="carrier") %>% 
  rename(total_flights_by_region=n.x) %>% 
  rename(total_carrier_flights=n.y)%>% 
  mutate(proportion = total_flights_by_region/total_carrier_flights)

ggplot(data= carrier_flights, aes(x=region, y=proportion)) + geom_bar(stat="identity") + facet_wrap(~carrier)

```
